### 2018-05-07 doc/paper.tex
Updating the L{F, k} lemma to actually be something useful.
This was simply the result of "sleeping" on it.

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\qed}{\(\blacksquare\)}
\newcommand{\cond}[1]{\operatorname{cond}\left(#1\right)}
\newcommand{\fl}[1]{\operatorname{fl}\left(#1\right)}
\newcommand{\eps}{\varepsilon}
\newcommand{\mach}{\mathbf{u}}
%% db == ``delta'' b
\newcommand{\db}[1]{
  \ifthenelse{\equal{#1}{1}}
             {\partial b}
             {\partial^{#1} b}
}
%% cdb == computed ``delta'' b
\newcommand{\cdb}[1]{
  \ifthenelse{\equal{#1}{1}}
             {\widehat{\partial b}}
             {\widehat{\partial^{#1} b}}
}

@@

In computer aided geometric design, polynomials are usually expressed in
Bernstein form. Polynomials in this form are usually evaluated by the
de Casteljau algorithm. This algorithm has a round-off error bound
which grows only linearly with degree, even though the number of
arithmetic operations grows quadratically. The Bernstein basis is
optimally suited (\cite{Farouki1987}, \cite{Delgado2015}, \cite{Mainar2005})
for polynomial evaluation; it is
typically more accurate than the monomial basis, for example in
Figure~\ref{fig:horner-inferior} evaluation via Horner's method produces
a jagged curve for points near a triple root, but the de Casteljau algorithm
produces a smooth curve. Nevertheless the de Casteljau
algorithm returns results arbitrarily less accurate than the working
precision \(\mach\) when evaluating \(p(s)\) is ill-conditioned.
The relative accuracy of the computed
evaluation with the de Casteljau algorithm (\texttt{DeCasteljau}) satisfies
(\cite{Mainar1999}) the following a priori bound:
\begin{equation}\label{de-casteljau-error}
  \frac{\left|p(s) - \mathtt{DeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq
  \cond{p, s} \times \mathcal{O}(\mach).
\end{equation}
In the right-hand side of this inequality, \(\mach\) is the computing
precision and the condition number \(\cond{p, s} \geq 1\) only depends
on \(s\) and the Bernstein coefficients of \(p\) --- its expression will
be given further.

@@

To address ill-conditioned problems, error-free transformations (EFT) can
be applied in \textit{compensated algorithms} to account for roundoff.
Error-free transformations were studied in great detail in \cite{Ogita2005}
and open a large number of applications.
In \cite{langlois_et_al:DSP:2006:442}, a compensated Horner's algorithm was
described to evaluate a polynomial in the monomial basis. In \cite{Jiang2010},
a similar method was described to perform a compensated version of the de
Casteljau algorithm. In both cases, the \(\cond{p, s}\) factor is moved
from \(\mach\) to \(\mach^2\) and the computed value is as accurate
as if the computations were done in twice the working precision. For example,
the compensated de Casteljau algorithm (\texttt{CompDeCasteljau}) satisfies
\begin{equation}\label{de-casteljau-2-error}
  \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \mathcal{O}\left(\mach^2\right).
\end{equation}
For problems with \(\cond{p, s} < 1 / \mach^2\), the relative error
is \(\mach\), i.e. accurate to full precision, aside from rounding to the
nearest floating point number. Figure~\ref{fig:jlcs-10} shows this shift
in relative error from \texttt{DeCasteljau} to \texttt{CompDeCasteljau}.

@@

As in~\eqref{de-casteljau-error}, the accuracy of the compensated
result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
polynomial evaluations. For example, as the condition number grows in
Figure~\ref{fig:jlcs-10}, some points have relative error exactly equal to
\(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
a complete failure to evaluate the order of magnitude of \(p(s)\). For
root-finding problems \(\mathtt{CompDeCasteljau}(p, s) = 0\) when
\(p(s) \neq 0\) can cause premature convergence and incorrect results.
We describe how to defer rounding into progressively
smaller error groups and improve the accuracy of the computed result by a
factor of \(\mach\) for every error group added. So we derive
\texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
that satisfies the following a prior bound for any arbitrary integer \(K\):
\begin{equation}
  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \mathcal{O}\left(\mach^K\right).
\end{equation}
This means that the computed value with \texttt{CompDeCasteljauK} is now
as accurate as the result of the de Casteljau algorithm performed in
\(K\) times the working precision with a final rounding back to the
working precision.

@@

Unfortunately, Figure~\ref{fig:compensated-insufficient} shows how
\texttt{CompDeCasteljau} starts to break down in a region of
high condition number (caused by a multiple root with multiplicity
higher than two). For example, the point
\(s = \frac{1}{2} + 1001\mach\)
--- which is in the plotted region \(\left|s - \frac{1}{2}\right|
\leq \frac{3}{2} \cdot 10^{-11}\) --- evaluates to exactly \(0\) when
it should be \(\mathcal{O}\left(\mach^3\right)\). As shown in
Table~\ref{tab:exact-computation}, the breakdown occurs because
\(\widehat{b}_0^{(0)} = -\cdb{1}_0^{(0)} = \mach / 16\).

@@

\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
requires
\begin{equation}
(15K^2 + 11K - 34)T_n + 6K^2 - 11K + 11 =
\mathcal{O}\left(n^2 K^2\right)
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\). Another way
to lower the total flop count is to just use
\(\widehat{b}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}\)
instead of \texttt{SumK}. This will reduce the total by
\(6(K - 1)^2\) flops, however the results produced are (empirically)
identical. This makes sense: the whole point of \texttt{SumK}
is to filter errors in a summation so that the final operation produces
a sum of the form \(v_1 \oplus \cdots \oplus v_K\) where each
term is smaller than the previous by a factor of \(\mach\). This
property is already satisfied for the \(\cdb{F}_0^{(0)}\) so in
practice the \(K\)-compensated summation is likely not needed.

@@

\begin{lemma}\label{lemma:ell-tilde}
The local errors \(\widetilde{\ell}_{F, j}\) satisfy:
\begin{align}
\widetilde{\ell}_{1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right) \\
\widetilde{\ell}_{F + 1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\cdb{F}_j^{(k + 1)}\right| +
  s \left|\cdb{F}_{j + 1}^{(k + 1)}\right|\right) +
  \gamma_{5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
  \text{ for } F \geq 1.
\end{align}
\end{lemma}

As we'll see soon (Lemma~\ref{lemma:k-order}), putting a bound on
sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
be necessary to get an overall bound on the relative error for
\texttt{CompDeCasteljauK}. So we define
\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) and the
related
\(D_{0, k} \coloneqq \sum_{j = 0}^k \left|\widehat{b}_j^{(k)}\right|
B_{j, k}(s)\) and
\(D_{F, k} \coloneqq \sum_{j = 0}^k \left|\cdb{F}_j^{(k)}\right| B_{j, k}(s)\).

\begin{lemma}\label{lemma:L-and-D-bounds}
For \(s \in \left[0, 1\right]\), the Bernstein-type sums defined above
satisfy the following bounds:
\begin{align}
D_{0, n - k} &\leq \left(1 + \gamma_{3k}\right) \cdot \widetilde{p}(s) \\
L_{1, n - k} &\leq \gamma_{3k} \cdot \widetilde{p}(s) \\
D_{F + 1, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
  \cdots \gamma_{5F} \cdot \left(1 + \gamma_{3k + 5F + 2}\right)
  \cdot \widetilde{p}(s) \\
L_{F + 2, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
  \cdots \gamma_{5F} \cdot \gamma_{3k + 5F + 2} \cdot \widetilde{p}(s).
\label{L-sum-bound}
\end{align}
In particular, this means that
\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\) and
\(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\).
\end{lemma}

See the appendix (Section~\ref{sec:appendix-proof-details}) for details.

@@

\begin{proof}
As in \eqref{matrix-de-casteljau}, we can express the compensated
de Casteljau algorithm as
\begin{equation}
\db{F}^{(k)} = U_{k + 1} \db{F}^{(k + 1)} + \ell_{F}^{(k)}
\Longrightarrow \db{F}^{(0)} = \sum_{k = 0}^{n - 1}
U_1 \cdots U_k \ell_F^{(k)} = \sum_{k = 0}^{n - 1}
\left[\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\right].
\end{equation}
For the inexact equivalent of these things, first note that
\(\widehat{r} = (1 - s)(1 + \delta)\). Due to this,
we put the \(\widehat{r}\) term at the end of each update step to reduce
the amount of round-off:
\begin{align}
  \cdb{F}_j^{(k)} &=
  \widehat{\ell}_{F, j}^{(k)} \oplus
  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right) \\
&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(k)} &=
  U_{k + 1} \cdb{F}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(0)} &=
  \sum_{k = 0}^{n - 1}
  U_1 \cdots U_k \widehat{\ell}_F^{(k)} (1 + \theta_{3k + 2})
  = \sum_{k = 0}^{n - 1}
  \left[\sum_{j = 0}^k \widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})
    B_{j, k}(s)\right].
\end{align}
Since
\begin{equation}
\db{F + 1}_0^{(0)} = \db{F}_0^{(0)} - \cdb{F}_0^{(0)} = \sum_{k = 0}^{n - 1}
\sum_{j = 0}^k \left(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right) B_{j, k}(s)
\end{equation}
it's useful to put a bound on \(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\). Via
\begin{align}
\widehat{\ell}_{F, j}^{(k)} &= e_1 \oplus \cdots \oplus e_{5F - 2} \oplus
\left(\rho \otimes \cdb{F - 1}_j^{(k + 1)}\right) \\
&= e_1\left(1 + \theta_{5F - 2}\right) + \cdots +
e_{5F - 2}\left(1 + \theta_2\right) +
\rho \cdot \cdb{F - 1}_j^{(k + 1)} \left(1 + \theta_2\right)
\end{align}
we see that
\begin{equation}
\left|\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right| \leq
\gamma_{3k + 5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
\Longrightarrow
\left|\db{F + 1}_0^{(0)}\right| \leq \sum_{k = 0}^{n - 1}
\gamma_{3k + 5F} \sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s).
\end{equation}
Since Lemma~\ref{lemma:L-and-D-bounds} (in particular \eqref{L-sum-bound})
tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
= L_{F, k} = \mathcal{O}\left(\mach^F\right)\) ...
\end{proof}

\begin{theorem}
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
      \right] +
      \left[\alpha(n) \mach^K + \mathcal{O}\left(
      \mach^{K + 1}\right)\right] \cond{p, s}.
  \end{equation}
\end{theorem}

\begin{proof}
Since
\begin{equation}
\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)},
\end{equation}
applying Theorem~\ref{thm:sum-k} tells us that
\begin{equation}\label{sum-k-applied}
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right|.
\end{equation}
Since
\begin{equation}
p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
= \cdots
= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
\end{equation}
we have
\begin{gather}
\left|\sum_{F = 0}^{K - 1} \cdb{F}_0^{(0)}\right|
\leq \left|p(s)\right| + \left|\db{K}_0^{(0)}\right| \quad \text{and} \\
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \leq
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\left|\db{K}_0^{(0)}\right| \label{triangle-ps}.
\end{gather}
Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
\mathcal{O}\left(\mach^F\right) \widetilde{p}(s)\), hence
\begin{align}
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| &\leq
\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
\gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
&\leq
\gamma_{2n - 2}^K \left[\left|p(s)\right| +
  \mathcal{O}\left(\mach\right) \widetilde{p}(s)\right] +
\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s).
\end{align}
Combining this with \eqref{sum-k-applied} and \eqref{triangle-ps}, we
see
\begin{align}
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| &\leq
\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
\left|\db{K}_0^{(0)}\right| +
\mathcal{O}\left(\mach^{K + 1}\right) \widetilde{p}(s) \\
&\leq
\left[\mach + \mathcal{O}\left(\mach^2\right)\right] \left|p(s)\right| +
\left[\alpha(n) \mach^K + \mathcal{O}\left(\mach^{K + 1}\right) \right]
\widetilde{p}(s).
\end{align}
Dividing this by \(\left|p(s)\right|\), we have our result.
\end{proof}

>>>>

\renewcommand{\headrulewidth}{0pt}
\renewcommand{\qed}{\(\blacksquare\)}
\newcommand{\cond}[1]{\operatorname{cond}\left(#1\right)}
\newcommand{\fl}[1]{\operatorname{fl}\left(#1\right)}
\newcommand{\bigO}[1]{\mathcal{O}\left(#1\right)}
\newcommand{\mach}{\mathbf{u}}
%% db == ``delta'' b
\newcommand{\db}[1]{
  \ifthenelse{\equal{#1}{1}}
             {\partial b}
             {\partial^{#1} b}
}
%% cdb == computed ``delta'' b
\newcommand{\cdb}[1]{
  \ifthenelse{\equal{#1}{1}}
             {\widehat{\partial b}}
             {\widehat{\partial^{#1} b}}
}

@@

In computer aided geometric design, polynomials are usually expressed in
Bernstein form. Polynomials in this form are usually evaluated by the
de Casteljau algorithm. This algorithm has a round-off error bound
which grows only linearly with degree, even though the number of
arithmetic operations grows quadratically. The Bernstein basis is
optimally suited (\cite{Farouki1987}, \cite{Delgado2015}, \cite{Mainar2005})
for polynomial evaluation; it is
typically more accurate than the monomial basis, for example in
Figure~\ref{fig:horner-inferior} evaluation via Horner's method produces
a jagged curve for points near a triple root, but the de Casteljau algorithm
produces a smooth curve. Nevertheless the de Casteljau
algorithm returns results arbitrarily less accurate than the working
precision \(\mach\) when evaluating \(p(s)\) is ill-conditioned.
The relative accuracy of the computed
evaluation with the de Casteljau algorithm (\texttt{DeCasteljau}) satisfies
(\cite{Mainar1999}) the following a priori bound:
\begin{equation}\label{de-casteljau-error}
  \frac{\left|p(s) - \mathtt{DeCasteljau}(p, s)\right|}{\left|p(s)\right|} \leq
  \cond{p, s} \times \bigO{\mach}.
\end{equation}
In the right-hand side of this inequality, \(\mach\) is the computing
precision and the condition number \(\cond{p, s} \geq 1\) only depends
on \(s\) and the Bernstein coefficients of \(p\) --- its expression will
be given further.

@@

To address ill-conditioned problems, error-free transformations (EFT) can
be applied in \textit{compensated algorithms} to account for roundoff.
Error-free transformations were studied in great detail in \cite{Ogita2005}
and open a large number of applications.
In \cite{langlois_et_al:DSP:2006:442}, a compensated Horner's algorithm was
described to evaluate a polynomial in the monomial basis. In \cite{Jiang2010},
a similar method was described to perform a compensated version of the de
Casteljau algorithm. In both cases, the \(\cond{p, s}\) factor is moved
from \(\mach\) to \(\mach^2\) and the computed value is as accurate
as if the computations were done in twice the working precision. For example,
the compensated de Casteljau algorithm (\texttt{CompDeCasteljau}) satisfies
\begin{equation}\label{de-casteljau-2-error}
  \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \bigO{\mach^2}.
\end{equation}
For problems with \(\cond{p, s} < 1 / \mach^2\), the relative error
is \(\mach\), i.e. accurate to full precision, aside from rounding to the
nearest floating point number. Figure~\ref{fig:jlcs-10} shows this shift
in relative error from \texttt{DeCasteljau} to \texttt{CompDeCasteljau}.

@@

As in~\eqref{de-casteljau-error}, the accuracy of the compensated
result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
polynomial evaluations. For example, as the condition number grows in
Figure~\ref{fig:jlcs-10}, some points have relative error exactly equal to
\(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
a complete failure to evaluate the order of magnitude of \(p(s)\). For
root-finding problems \(\mathtt{CompDeCasteljau}(p, s) = 0\) when
\(p(s) \neq 0\) can cause premature convergence and incorrect results.
We describe how to defer rounding into progressively
smaller error groups and improve the accuracy of the computed result by a
factor of \(\mach\) for every error group added. So we derive
\texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
that satisfies the following a prior bound for any arbitrary integer \(K\):
\begin{equation}
  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \bigO{\mach^K}.
\end{equation}
This means that the computed value with \texttt{CompDeCasteljauK} is now
as accurate as the result of the de Casteljau algorithm performed in
\(K\) times the working precision with a final rounding back to the
working precision.

@@

Unfortunately, Figure~\ref{fig:compensated-insufficient} shows how
\texttt{CompDeCasteljau} starts to break down in a region of
high condition number (caused by a multiple root with multiplicity
higher than two). For example, the point
\(s = \frac{1}{2} + 1001\mach\)
--- which is in the plotted region \(\left|s - \frac{1}{2}\right|
\leq \frac{3}{2} \cdot 10^{-11}\) --- evaluates to exactly \(0\) when
it should be \(\bigO{\mach^3}\). As shown in
Table~\ref{tab:exact-computation}, the breakdown occurs because
\(\widehat{b}_0^{(0)} = -\cdb{1}_0^{(0)} = \mach / 16\).

@@

\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
requires
\begin{equation}
(15K^2 + 11K - 34)T_n + 6K^2 - 11K + 11 =
\bigO{n^2 K^2}
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\). Another way
to lower the total flop count is to just use
\(\widehat{b}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}\)
instead of \texttt{SumK}. This will reduce the total by
\(6(K - 1)^2\) flops, however the results produced are (empirically)
identical. This makes sense: the whole point of \texttt{SumK}
is to filter errors in a summation so that the final operation produces
a sum of the form \(v_1 \oplus \cdots \oplus v_K\) where each
term is smaller than the previous by a factor of \(\mach\). This
property is already satisfied for the \(\cdb{F}_0^{(0)}\) so in
practice the \(K\)-compensated summation is likely not needed.

@@

\begin{lemma}\label{lemma:ell-tilde}
The local errors \(\widetilde{\ell}_{F, j}\) satisfy:
\begin{align}
\widetilde{\ell}_{1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right)
  \label{ell-tilde-1} \\
\widetilde{\ell}_{F + 1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\cdb{F}_j^{(k + 1)}\right| +
  s \left|\cdb{F}_{j + 1}^{(k + 1)}\right|\right) +
  \gamma_{5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
  \text{ for } F \geq 1.
\end{align}
\end{lemma}

As we'll see soon (Lemma~\ref{lemma:k-order}), putting a bound on
sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
be necessary to get an overall bound on the relative error for
\texttt{CompDeCasteljauK}, so we define
\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\).

\begin{lemma}\label{lemma:L-and-D-bounds}
For \(s \in \left[0, 1\right]\), the Bernstein-type sum defined above
satisfies the following bound:
\begin{equation}\label{L-sum-bound}
L_{F, n - k} \leq \left[\left(3^F \binom{k}{F - 1} + \bigO{k^{F - 1}}\right)
  \mach^F + \bigO{\mach^{F + 1}}\right] \cdot \widetilde{p}(s).
\end{equation}
In particular, this means that
\(L_{F, k} = \bigO{(3 \mach)^F} \cdot \widetilde{p}(s)\).
\end{lemma}

See the appendix (Section~\ref{sec:appendix-proof-details}) for details on
proving Lemma~\ref{lemma:ell-tilde} and Lemma~\ref{lemma:L-and-D-bounds}.

@@

\begin{proof}
As in \eqref{matrix-de-casteljau}, we can express the compensated
de Casteljau algorithm as
\begin{equation}
\db{F}^{(k)} = U_{k + 1} \db{F}^{(k + 1)} + \ell_{F}^{(k)}
\Longrightarrow \db{F}^{(0)} = \sum_{k = 0}^{n - 1}
U_1 \cdots U_k \ell_F^{(k)} = \sum_{k = 0}^{n - 1}
\left[\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\right].
\end{equation}
For the inexact equivalent of these things, first note that
\(\widehat{r} = (1 - s)(1 + \delta)\). Due to this,
we put the \(\widehat{r}\) term at the end of each update step to reduce
the amount of round-off:
\begin{align}
  \cdb{F}_j^{(k)} &=
  \widehat{\ell}_{F, j}^{(k)} \oplus
  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right) \\
&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(k)} &=
  U_{k + 1} \cdb{F}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(0)} &=
  \sum_{k = 0}^{n - 1}
  U_1 \cdots U_k \widehat{\ell}_F^{(k)} (1 + \theta_{3k + 2})
  = \sum_{k = 0}^{n - 1}
  \left[\sum_{j = 0}^k \widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})
    B_{j, k}(s)\right].
\end{align}
Since
\begin{equation}
\db{F + 1}_0^{(0)} = \db{F}_0^{(0)} - \cdb{F}_0^{(0)} = \sum_{k = 0}^{n - 1}
\sum_{j = 0}^k \left(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right) B_{j, k}(s)
\end{equation}
it's useful to put a bound on \(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\). Via
\begin{align}
\widehat{\ell}_{F, j}^{(k)} &= e_1 \oplus \cdots \oplus e_{5F - 2} \oplus
\left(\rho \otimes \cdb{F - 1}_j^{(k + 1)}\right) \\
&= e_1\left(1 + \theta_{5F - 2}\right) + \cdots +
e_{5F - 2}\left(1 + \theta_2\right) +
\rho \cdot \cdb{F - 1}_j^{(k + 1)} \left(1 + \theta_2\right)
\end{align}
we see that
\begin{equation}
\left|\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right| \leq
\gamma_{3k + 5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
\Longrightarrow
\left|\db{F + 1}_0^{(0)}\right| \leq \sum_{k = 0}^{n - 1}
\gamma_{3k + 5F} \sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s).
\end{equation}
Since Lemma~\ref{lemma:L-and-D-bounds} (in particular \eqref{L-sum-bound})
tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
= L_{F, k} = \bigO{\mach^F}\) ...
\end{proof}

\begin{theorem}
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
      \left|p(s)\right|} \leq \left[\mach + \bigO{\mach^2}
      \right] +
      \left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}}\right] \cond{p, s}.
  \end{equation}
\end{theorem}

\begin{proof}
Since
\begin{equation}
\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)},
\end{equation}
applying Theorem~\ref{thm:sum-k} tells us that
\begin{equation}\label{sum-k-applied}
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right|.
\end{equation}
Since
\begin{equation}
p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
= \cdots
= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
\end{equation}
we have
\begin{gather}
\left|\sum_{F = 0}^{K - 1} \cdb{F}_0^{(0)}\right|
\leq \left|p(s)\right| + \left|\db{K}_0^{(0)}\right| \quad \text{and} \\
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \leq
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\left|\db{K}_0^{(0)}\right| \label{triangle-ps}.
\end{gather}
Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
\bigO{\mach^F} \widetilde{p}(s)\), hence
\begin{align}
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| &\leq
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
\gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
&\leq
\gamma_{2n - 2}^K \left[\left|p(s)\right| +
  \bigO{\mach} \widetilde{p}(s)\right] +
\bigO{\mach^{K + 1}} \widetilde{p}(s).
\end{align}
Combining this with \eqref{sum-k-applied} and \eqref{triangle-ps}, we
see
\begin{align}
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| &\leq
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left|\db{K}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
&\leq
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left[\alpha(n) \mach^K + \bigO{\mach^{K + 1}} \right]
\widetilde{p}(s).
\end{align}
Dividing this by \(\left|p(s)\right|\), we have our result.
\end{proof}
